\documentclass{report} %skal vi legge til [a4paper] og endre til report? 

\usepackage[a4paper]{geometry}
\usepackage[backend=bibtex, style=verbose-trad2]{biblatex}
\usepackage[utf8]{inputenc}
\usepackage{comment}
\bibliography{kilder}

\title{INF226 Obligatory assignment}
\date{2016-10-02}
\author{Håvar Eggereide and Syver Storm-Furru}

\begin{document}
\maketitle
\pagenumbering{gobble}
\newpage
\pagenumbering{arabic}
\tableofcontents
\newpage
\chapter{Introduction}

  For our obligatory assignment we were tasked with analysing OpenMRS, a medical
  patient journal system. The analysis is performed both using static and
  dynamic code analysis tools (HPE Fortify and FindBugs for static analysis and
  OWASP ZAP for dynamic), as well as a thorough manual run-through of the installation
  process and software usage.

  \paragraph{The goals of software testing}

  It is now common practice  for a developer to adhere to a software architect's
  plan for the development of the software components, and many utilizes
  unit-testing as a supporting tool for developing code without errors. So what
  is then the benefit of having a separate plan and procedure for system testing
  from the view of security?

  It has been demonstrated with many grim examples during this course and in the
  curriculum texts, the importance of security and the heavy consequences to the
  whole enterprise if it is neglected. The issue is of such a big scope that
  approaching it as just another design or implementation issue fails to
  appreciate the complexity it presents.

  The goal of testing the security as a separate process is to try to map as
  much as possible of the potential security issues and security considerations
  that has to be kept in mind walking thru the process of development. To
  accomplish this it is necessary to approach the software and the architecture
  from the point of view of a agent with malicious intent.

  In the end it is  most likely not possible to make the software completely
  immune to security faults. It's a natural part of security testing to find
  problems where a full repair might not be achievable, it is then also the
  responsibility of the security testing process to develop mitigation plans for
  these vulnerabilities.

  So the goal of security testing is to find, repair and prepare for security
  vulnerabilities.
  
  \paragraph{Static and dynamic analysis}

  Static and dynamic analysis are methods for checking the software for vulnerabilities.
  They are both done using software tools that help the user analyze. Static
  analysis is performed before any code is run, and checks the code base for
  vulnerable code that can be abused. It checks different runtime behaviours,
  both standard and non-standard, in order to find potential backdoors, weak
  code and other vulnerabilities. Dynamic analysis on the other hand is done
  while the application is running, and works a lot like how an actual attacker
  would target the application, performing such things as XSS and SQL incjection
  attacks, as well as monitoring the system runtime, for example performance and
  HTTP requests \autocite[]{VERACODE}.

\newpage
\section{Software involved}
  
\subsection{OpenMRS}

  OpenMRS is, according to the website, the ``world's leading open source
  enterprise electronic medical record system platform'' \autocite[]{OpenMRSAbout}. It is used in
  hospitals and medical facilities all over the world, for example in Nigeria, South Africa, India
  and the United States, and is supported by many different governments, NGOs,
  and both for- and non-profit organisations. The software has a stated goal of
  being usable with no programming knowledge, and to be a common platform for
  which medical informatics efforts in developing can be built.

  \paragraph{Technical Specifications}

  OpenMRS is a client-server platform, with a web front end. It is programmed in
  Java 7, using Tomcat 6 or 7 as the server framework, and MySQL 5.6 as the
  database backend. It also exposes a programming API to users, and is modular
  and extendable.

  

  \paragraph{Usage}

  OpenMRS is operated through a web browser, and is centered around a main dashboard that,
  with just the base installation, will give you access to patient journals, as well as
  showing currently active patients. The system also lets you categorize patients/other registered
  in different persons in different categories, easing the process of generating statical analyses.
  The system also lets your register and book future meetings with patients.

\subsection{HPE Fortify}

  Fortify is a code security tool suite, developed by Hewlett-Packard
  Enterprise (HPE). It aims to ``make application security a natural part of the new
  SDLC, enabling time to market by building security in''\autocite[]{Fortify}.
  It contains such tools as WebInspect, a dynamic code analysis tool, and the
  Fortify Static Code Analysis tool. Fortify is a proprietary solution, but is
  available with an academic license for free.

  We ran the Audit Workbench version bundled with Fortify version 16.10.

\paragraph{Audit Workbench}

  Audit Workbench is the tool used to organise the the output of HPE's static
  code analysis software contained in the Fortify package. It is a GUI
  application built on top of Eclipse, specially designed for organising and
  presenting output for the HPE tools. 
  
  The installation process for Audit Workbench was straight-forward, but running
  the program required changing variables for the Eclipse backend, without
  information about how this is done readily available. The software also was a
  large RAM consumer, needing 5 gigabytes of RAM to analyze a relatively large
  project (OpenMRS).

\subsection{FindBugs}

FindBugs is a static analysis tool for scanning java code looking for potential errors in the
implementation. It is distributed under Lesser GNU Public License. The project
originated from the University of Maryland.\autocite[]{FindBugs} It runs on java
1.7 ether initiated from a terminal or you may use a simple GUI which makes it
easier to get familiar with the program. 

We ran FindBugs 3.0.1 for our tests.
 

\subsection{OWASP ZAP}

ZAP is a dynamic analysis tool developed by OWASP, the Open Web Security
Project. The software acts as a proxy between the host computer and a web
application, performing different types of automatic scans, as well as having
tools for manual searches for security vulnerabilities\autocite[]{ZAP}.

We ran ZAP version 2.5.0 for our tests.

\subsection{ThreadSafe} 

``ThreadSafe is a static analysis tool for finding concurrency bugs and
potential performance issues in Java programs.''
 \autocite[]{THREADSAFE}

The program is developed by Contemplate ltd. aiming to help developers discover
concurrency issues in the code-base prior to the deploying of the application.
Concurrency is a complicated issue, having the correct understanding of
what it is and how it works is not easy. Then the next step
of implementing it correctly can give rise to more complicated issues which are
even harder to discover. For the developer to have a tool to aid in working with
concurrency is of high value. Should a issue arise during runtime the
ramifications, for a system like OpenMRS which potentially need to be operating
24 hours a day, could possibly be disruption of the integrity of the database
and halt operations.

ThreadSafe is not open source and it is necessary to invest in a license if
there is a need to integrate the application in the SDLC. There is however a
free trial of 14 days available.
%  LocalWords:  HPE

\chapter{Setup and Configuration}
\subsection{OpenMRS}
  The setup process of OpenMRS is quite involved and time consuming when
  attempting to do so on a personal computer, requiring both Tomcat, Java and
  MySQL to be setup. The official documentation is useful, but different parts
  of it uses different versions of e.g. Tomcat, so it can be confusing.
  It also provides install instructions for Windows and Linux distributions
  with Aptitude, but not for OSX or other Linux distributions. 

  As mentioned, OpenMRS doesn't run on the newest version of MySQL (at the time
  of writing MySQL 5.7), and the install instructions do not mention this. The
  process of figuring this out, and of removing and reinstalling a previous
  version of MySQL, proved to be a lengthy detour on an already long road.
  The instructions are also not very specific when noting which files you need
  to run OpenMRS in Tomcat, whether it is the source code, which was difficult
  to build and only return a test suite on a normal compile, a readily packaged
  complete install (which did not work properly), or a .war file that should be
  uploaded to the Tomcat server.

  We first attempted setting up OpenMRS on OSX 10.11, but ran into problems when
  trying to install the correct version of MySQL, and therefore retried in an
  empty virtual machine running Linux (tested with both Lubuntu and Kali Linux).
  Following the install instructions were a lot easier when running Debian based
  distributions containing Aptitude, but we still had to find and install a
  previous version of MySQL. 

  Once everything was installed OpenMRS had some extra setup that was required,
  done through a web interface. This was mostly easy once the correct version of
  MySQL was in place. You were also given a first username and password that
  was, respectively, `admin' and `Admin123'.


\subsection{Tools}
\paragraph{Audit Workbench}

Audit Workbench was set up by downloading the HPE Fortify software suite, which
contains installers for different Fortify software. The installer for Audit
Workbench was run, without any issues. Setting up the Audit Workbench IDE
required changing the -Xmx setting in a deeply nested ``eclipse.ini'' file,
which required a bit of dedication to find. The reason for this is that scanning
a project requires more memory than is allocated to Audit Workbench by default.
After that, running Audit Workbench was pretty straight forward.

\paragraph{FindBugs}

The stand-alone application which is distributed from the projects main page was
easy to get started from the terminal(OSX-bash) without any setup other than the
first installation. Starting the initial scan was also straight forward. The
high number of errors reported was misleading and proved a cumbersome task. What
proved to be a better solution was using a FindBug-plugin for the IntelliJ IDEA.
This way IntelliJ made a build of the code base and then ran the scan on this
build. IntelliJ could reason about the classes which where not necessary
to scan, as per example all the unit tests. The code base we used for the scan was
cloned directly from github (https://github.com/openmrs/openmrs-core).

\paragraph{OWASP Zap}

Running ZAP requires ZAP to be installed in the same location as OpenMRS when
running OpenMRS locally. Therefore, we set up a virtual machine to run both ZAP,
the Tomcat server and the MySQL server. The virtual machine used ran the the
linux distribution Kali Linux, which is a specialized distro for penetration
testing. It included OWASP ZAP, and we therefore did not need to set up the
application itself. However, fully utilizing ZAP required using an addon for our
browser (Firefox). The addon had an out-of-date signature, and thus needed
Firefox to run addons without valid signatures. After changing that setting for
Firefox and running the setup wizard for ZAP inside Firefox, it was ready to be used.

\paragraph{ThreadSafe} % Setup and Configuration

ThreadSafe has three install options, two which are available in the trial
version. These options are: A .jar file which is executable from the terminal
taking input-arguments and a setup file for each project, plugin for a static
analyzing tool called SonarQube and a plugin for the Eclipse IDE.

As we were using a free trial we had to be limited to ether using the terminal
program or the Eclipse plugin, which were the included options in the trial. As
setting up a project configuration file for the terminal application seemed to
open the possibility of misconfiguration, using the Eclipse plugin seemed the
best alternative. Installing the plugin from the .zip file was easy, the only
other thing necessary was to order a trial license and past this license in the
Preferences -> ThreadSafe. We then let Eclipse build the Maven .pom file from
the OpenMRS source-code and then everything was ready for scanning. 

\section{Preliminary results}

\subsection{Visible potential issues}
  During the install process we were on the lookout for potential vulnerabilities that
  were visible to us. Two things stood out:
  1) The program was set up with a static username and password (resepctively `admin'
  and `Admin123'), and we were never prompted to change this. This means that if the system
  installer doesn't notice this is the case, the is a user with administrator priviledges
  available to anyone, which is pretty bad.
  2) The system didn't work with the newest version of MySQL. The latest version it can run (newest 5.6 version)
  has several vulnerabilities, some very serious \autocite[]{CVEDetails}. While some of these
  also are present in newer versions of MySQL, others are not, making the database slightly more vulnerable.

\chapter{Tests of the software tools}

  After we set everything up, we ran a test run of each analysis tool, to make sure
  everything was functioning properly and to find any glaring issues, if any.

\paragraph{Dynamic analysis of the OpenMRS Web GUI}
  
  The initial Quick Start scan of our test setup of the OpenMRS service only had
  the login front page to crawl. This yielded very few results considering this
  is just one page. It found some information about jQuery, ZAP also warned
  about the implementation of the cookie.
  
\paragraph{Audit Workbench report}  

  The first scan yielded a few messages of high concern. The OpenMRS core
  includes unit-test which gives quite a few erroneous warnings concerning
  security risks and more general bad coding practices. 

  There was an error concerning how the cookie was coded which may be
  interesting to look more into.

\paragraph{FindBugs code scan}

  Running the scan seems easy and most of the work is obviously analysing the
  results. The reports might be a bit hard to navigate and there was some
  trouble generating a html report that potentially is more readable. Again the
  unit-test generated false positives.

\paragraph{ThreadSafe} % Test of the software tools

The initial scan gave a few possible errors, all of them concerning ether the
OpenMRS-API or the OpenMRS-WEB which might make sense as these expose surface to
external modules. The scan also gave some false positives referring to
unit-test.
  
%  LocalWords:  SDLC SSL
  \newpage

\section{Findings}

\subsection{Raw data}

\paragraph{Fortify}

  Audit Workbench found 60 critical vulnerabilities, as well as 115 issues
  considered highly vulnerable. Following is a list of the categories designated
  by Audit Workbench for these vulnerabilities, as well as the number of
  occurrences of each vulnerability and a short explanation.

  The critical vulnerabilities were in the
  following categories:
  
  \begin{itemize}
  \item Command injections (4): The program tries to call an external command
    without validating where the command is from.
  \item Insecure SSL (3): Uses SSL certificates are based on default system
    Certificate Authorities, which may be compromised, and sometimes are.
  \item Hard coded passwords (3): Lets an attacker log in if he or she has access
    to the source code, which in this case means anyone.
  \item Path manipulation (2): The program accesses a file with a file path that
    can be chosen by the attacker.
  \item Privacy violation (49): The program mishandles confidential information
    in some way or another, for example by writing it to disk.
  \end{itemize}

  The vulnerabilities rated high are as follows:

  \begin{itemize}
  \item Header manipulation with cookies (1): The program doesn't validate data
    in an HTTP request, so cookies can be tampered or spoofed.
  \item Hard coded encryption key (3): Same issue as with hard coded passwords.
  \item Log forging (12): The program doesn't validate inputs that get written
    to the log, meaning malicious data could be entered.
  \item Empty passwords (6): The program assigns empty passwords, making it much
    easier to break into the application.
  \item Hard coded passwords (25): Same issue as with the critical hard coded passwords.
  \item Passwords in Configuration File (36): Passwords are stored in plain-text
    in a configuration file, meaning an attacker can get the password by getting
    the file.
  \item Privacy violation (10): Same issue as in the critical privacy violations.
  \item Privacy violation from heap inspection (3): The program stores
    confidential information in strings, which are immutable and thus stored in
    memory until the JVM garbage collector is run (which may not occur for a
    while). 
  \item Race Conditions from Singleton member field (19): The program uses a
    singleton class which takes concurrent connections, meaning data can be
    shared between users.
  \end{itemize}
      
\paragraph{Configuration Files}  

\paragraph{FindBugs}

  The first scan detected 1058 potential bugs. Some of them are from test classes

%  LocalWords:  JVM
  and will not be mentioned here. The report generated from the FindBugs GUI was
  very uninformative and hard to navigate. We found a plugin to IntelliJ IDEA
  that used Findbugs as the back-end and produced a more informative rapport
  where it was easy to exclude the testbase as IntelliJ had generated a build of the
  codebase giving at a sense of which code was where. In this new rapport we
  had a total of 296 warnings and a separation between what FindBugs was very
  certain was a bug and what maybe could be a bug. In the following text we have
  made a presentation of the bugs FindBugs is most certain is a potential
  problem.


Correctness
  \begin{itemize}
    \item Masked Field:

    Class defines field that masks field in superclass
    org.openmrs.web.filter.StartupFilter:
    This error is mostly harmless. It may have potential for generating
    errors in later updates to the code. The field is marked as
    protected in the superclass so it is excessive in the child and
    probably should be removed.

    \item Masked Field:

    Field UpdateFilter.log masks field in superclass
    org.openmrs.web.filter.StartupFilter:
    Same as above.
  \end{itemize}
Bad practice
  \begin{itemize}
    \item Dropped or ignored exception:

      This method might ignore an exception. In general, exceptions
      should be handled or reported in some way, or they should be thrown
      out of the method.

      This error occurred in the following classes:

      \begin{itemize}
        \item ServiceContext.setModuleService
        \item WorkflowCollectionEditor.setAsText
        \item OpenmrsUtil.parse
        \item BooleanConceptChangeSet.getInt
        \item BooleanConceptChangeSet.createConcept
      \end{itemize}
    \item Initialization circularity. 

      Class makes active use of subclass during initialization and might
      result in null value.

      Class:

      \begin{itemize}
        \item Operator 
      \end{itemize}
  \end{itemize}
Internationalization 
  \begin{itemize}
    \item Reliance on default encoding. 
      Found code that does a bit to string conversion while assuming
      the default platform encoding is suitable. This could make the code
      behave differently on different platforms.

      Classes:

      \begin{itemize}
        \item HL7ServiceImpl
        \item ModuleFileParser
        \item ModuleUtil
        \item TextHandler
        \item DatabaseUpdater
        \item HTTPClient
        \item OpenmrsUtil
        \item CreateCodedOrderFrequencyForDrugOrderFrequencyChangeset
        \item MigrateConceptReferenceTermChangeSet
        \item SourceMySqldiffFile
        \item Listener
        \item StartupFilter
        \item TestInstallUtil
      \end{itemize}
  \end{itemize}
Malicious code vulnerability
  \begin{itemize}
    \item Mutable static field. 
      Public static field isn't final and could be change by malicious
      code or by accident.

      Classes:

      \begin{itemize}
        \item SchedulerConstants
        \item OpenmrsConstants
        \item Security
      \end{itemize}
  \item Field is mutable array. 
    Same as above and field is referring to an array.

    Class:

    \begin{itemize}
      \item OpenmrsConstants
    \end{itemize}
  \end{itemize}
Multithreaded code vulnerability
  \begin{itemize}
    \item Unintended contention or possible deadlock due to locking on shared
      objects: 
      Synchronization on Boolean. There is a risk of syncing on a object
      belonging to another process since there is only two Boolean objects.
      Resulting in a deadlock

      Class:

      \begin{itemize}
        \item HL7InQueueProcessor
      \end{itemize}
    \item Unsynchronized Lazy Initialization: 
      Incorrect lazy initialization and update of static field. Static field
      is not completed and could be accessed by another process while it is
      still initializing which would lead to a multithreading bug.

      Class:
      \begin{itemize}
          \item OpenmrsClassLoader
      \end{itemize}
  \end{itemize}
Performance
  \begin{itemize}
    \item Questionable Boxing of primitive value. 
      Boxing a primitive to compare. It's more efficient to use static compare
      method on primitives.

      Class:
      \begin{itemize}
          \item ModuleFactory
      \end{itemize}
    \item Boxing/unboxing to parse a primitive. 
      A sting is boxed just to be unboxed for parsing. More
      efficient to use static parse*** method.

      Class:
      \begin{itemize}
        \item BaseHyphenatedIdentifierValidator
      \end{itemize}
  \end{itemize}
Security
  \begin{itemize}
    \item Potential SQL Problem. 
      Nonconstant string passed to execute or addBatch method on an SQL
      statement The method invokes the execute or addBatch method on an SQL
      statement with a String that seems to be dynamically generated.
      Consider using a prepared statement instead. It is more efficient
      and less vulnerable to SQL injection attacks. 

      Classes:
      \begin{itemize}
        \item ConceptValidatorChangeSet
        \item MigrateAllergiesChangeSet
      \end{itemize}
  \end{itemize}
Dodgy code
  \begin{itemize}
    \item Dead local store. This instruction assigns a value to a local variable, but the
      value is not read or used in any subsequent instruction. Often, this
      indicates an error, because the value computed is never used.
      Note that Sun's javac compiler often generates dead stores for final
      local variables. Because FindBugs is a bytecode-based tool, there is
      no easy way to eliminate these false positives.

      Class:
      \begin{itemize}

%  LocalWords:  multithreading
        \item PatientServiceImpl
      \end{itemize}
  \end{itemize}
\paragraph{OWASP ZAP}

A passive run of ZAP returns the following four possible vulnerabilities:

\begin{itemize}
 \item X-Frame-Options header is not included in the HTTP response to protect against
  'ClickJacking' attacks: Considered a medium vulnerability, this
  vulnerability can let unknown sites be loaded inside frames or iframes.
 \item A cookie has been set without the HttpOnly flag: A low vulnerability that
   makes it possible to access the cookie using Javascript. 
 \item Web browser XSS protection not enabled: A low vulnerability where the
   HTTP header X-XSS-Protection isn't set, so that the browser doesn't use it's
   own cross site scripting protection.
 \item X-Content-Type-Options Header missing: Some older browsers try to
   automatically figure out the content type if X-Content-Type-Options is not
   set to 'nosniff', making it possible to do such operations as cross-site
   scripting \autocite[]{XFrame}.
\end{itemize}

\paragraph{ThreadSafe} % Raw data

Excluding false positives from tests the scan gave 12 major and 1 minor
warnings. No Critical or ``Blocker'' warnings were generated. The major warnings
are as follows:

\begin{itemize}
\item Inconsistent synchronization of access to Field/a collection.

  StartupFilter, UpdateFilter.

\item Mixed synchronization of access to a field.

  OpenmrsClassLoader, OrderSetServiceImpl, UpdateFilter.
  
\item No lock held while iterating on a synchronized collection view.
  
  UpdateFilter.

\item Non atomic use of Get/Check/Put.

  WebModuleUtil.

\item Synchronizing on reusable objects.

  ModuleFactory, WebModuleUtil.

\item Unsafe iteration over synchronized collection.

  HL7InQueueProcessor.
  
\item Unsynchronized access to field from asynchronously invoked method.
  
  WebModuleUtil.

\end{itemize}

\subsection{Interpretations of the raw data}

\paragraph{Tests}
Some of the critical errors found by both Fortify and FindBugs were in test
classes, and are therefore not going to run while the application is deployed.
These bugs and vulnerabilities are therefore disregarded. This especially
applies to the Fortify results in the category Command Injection: The
application runs commands from external sources without validating them
beforehand.

\paragraph{External files}
The vulnerabilities concerning passwords in clear text found by Fortify are
mostly false positives, as they are found in localization files, and don't
contain real passwords (instead containing the translation of the noun
password). We will therefore disregard any vulnerability found in localization
files. However, some files containing clear text passwords are used for setting
up the server, and are thus relevant.

\paragraph{Privacy violations}
Many vulnerabilities found by Fortify were in the category Privacy Violations.
Considering the type of application OpenMRS is (it manages patient journals),
these are more important than for other applications, as patient journals
contain lots of personal and sensitive information such as (possibly) social security
numbers and medications used. Many of these files are about the the application
logging login information and search terms, and the security of the logger is therefore of
particular interest.

The other privacy violation needs the attacker to access the JVM memory, and
therefore requires the attacker to have root access. Since the program runs on a
server, if the attacker has root access you've probably lost your database and
similar serious problems, and thus problems requiring memory access aren't as
important. This might be more of a problem if the server is external (e.g.
hosted on Amazon AWS or CloudFlare), but then, it is up to Amazon or CloudFlare

%  LocalWords:  Javascript ClickJacking
to secure their servers.

\paragraph{Race conditions}
These errors arise from Servlet being a singleton. Since the application changes
member fields of this singleton, multiple request to get information can return


%  LocalWords:  iframes
%  LocalWords:  HttpOnly
wrong results (if the member field is changed during requests). 

\paragraph{Dynamic analysis}
The run of OWASP ZAP didn't find any critical or highly vulnerable issues,

%  LocalWords:  nosniff
indicating that the front end is relatively well secured. The ZAP issues were
related to HTTP headers (for example no SQL injection issues and direct XSS
issues). Since the critical issues were found by static analysis, we will focus
more on the server errors.

\subsection{Vulnerable external libraries/software}
In order to find vulnerabilities in libraries and software OpenMRS is based on
we looked at exploit-db.com, which is a database looking up vulnerabilities in
the CVE and OSVDB databases. We looked for exploits in Tomcat, JRE, Java, and
MySQL. We found the following relevant exploits:

\begin{itemize}
\item Tomcat: There are issues with the newest version of Tomcat (Tomcat 8) for
  OpenMRS, and we therefore looked for exploits in Tomcat 7. There are
  especially two exploits worth noting:
  \begin{itemize}
    \item EDB-ID 35011: Apache Tomcat 7.0.4 - 'sort' and 'orderBy' Parameters Cross-Site Scripting:
      Since Tomcat doesn't sanitize user inputs properly, cross-site scripting
      can be performed.
    \item EDB-ID 40450: Apache Tomcat 8/7/6 - Privilege escalation: The default
      Tomcat distributions available on Aptitude (Debian and Debian based distro
      package manager) as well as RPM (RedHat and RedHat derivate distro packagr
      manager) have a systemd config file that is insecure, run on boot, and can give
      root access to an attacker. The payload for this file can be delivered
      through a Tomcat shell in java web apps, or locally. This is a new exploit
      (found 10/10/2016) and thus unlikely to be patched on servers running OpenMRS.
  \end{itemize}
\item MySQL: EDB-ID 40678: MySQL/MariaDB/PerconaDB System user Privilege
  escalation / Race condition:
  Lets a local user access and run database operations above their own privilege
  level, and combined with other exploits, can give server root access to a user.
  
\item Java/JRE: OpenMRS runs on JRE 1.7, for which we couldn't find any relevant exploits.
\end{itemize}\autocite[]{exploitdb}

%  LocalWords:  XSS OWASP Servlet CloudFlare OpenMRS FindBugs
\chapter{Analysis, evaluation and recommendations}  

\section{Evaluation of external libraries}
\paragraph{MySQL}
Being able to access and edit the database is very dangerous in a system that
contains such sensitive data as patient journals.
As noted, the EDB-ID 40678 vulnerability in MySQL could be used to gain root
access to the database. It requires local access, but since OpenMRS can be used
for example in a hospital with a lot of local users of different privileges (for
example Medical Student, Laboratory Assistants and of course Doctors), the
vulnerability is very relevant to OpenMRS. 

\paragraph{Tomcat}
Since a lot of servers running OpenMRS are Linux servers running e.g. Ubuntu
(the official installation instructions list Windows and Ubuntu installation
steps), EDB-ID 40450 is a dangerous exploit. It can be exploited through a web
shell delivered through for example cross-site scripting or SQL injection, and
can give root access to the server, giving the attacker full control.

\section{Recommendations}

\paragraph{Updating source code}

\paragraph{Recommended software}
The install instructions for OpenMRS links the user to Tomcat 6, and with a
generic statement about how ``there are issues with Tomcat 8.''. In light of the
Tomcat exploit found on Exploit-DB, the user should not be using a version of Tomcat
from the package manager of the operating system if running a Linux server.
Downloading the newest version of Tomcat 7 from Apache's homepage is therefore
the reccomended course of action.

\paragraph{User management}
% Kopiert fra FlossManuals, bør omskrives
OpenMRS has three unchangeable user groups, Anonymous, Authenticated and System
Developer. Floss Manuals (which the official OpenMRS wiki refers to), there are
some specific recommendations for the privileges these users should recieve.
From FlossManuals: ``Anonymous privileges should be kept very restricted, since patient
information might otherwise be compromised. Privileges granted to the
Authenticated role are granted to anyone that logs into your system, no matter
what other role(s) they might be assigned. Granting privileges to the
Authenticated role is an easy way to grant privileges to all users of the
system. The System Developer role is automatically granted full access to the
system and should only be granted to system administrators.

Super users (system administrators) are automatically granted all privileges
in the system; therefore, you must be very careful to protect your system
administrator password.'' \autocite[]{FLOSS}
We recommend that users of the application follow these guidelines, as not
complying can lead to granting privileges to all users of the system, ergo more
easily letting attackers gain information not usually available to them.

\paragraph{Race conditions}
Of the different vulnerabilities, the ones concerning race conditions are 
the second largest group. Since race conditions are becoming an increasingly
more commmon cause of errors and can cause such errors as returning critical
data, to non-authorized users, these errors should be considered extra important
to fix.

\autocite[]{FLOSS}

\section{Summary}

\end{document}


